Multilocus Genotype And Mycelial Compatability Group Assessment

### 2017-05-04 doc/RMD/mlg-mcg.Rmd
Add trees to mlg-mcg debacle

cnames <- LETTERS[1:10]
cnames[c(7, 5, 1)] <- LETTERS[1:3]
the_communities <- data_frame(vertex = lay2$name, community = lay2$community) %>%
  mutate(community = case_when(
    .$community == 7 ~ "A",
    .$community == 5 ~ "B",
    .$community == 1 ~ "C",
    TRUE ~ as.character(.$community)
  )) %>%
  group_by(community) %>%
  mutate(comm = ifelse(n() > 10, paste("Community", community), "Other Communities (n < 10)")) %>%
  mutate(comm = comm) # It just so happens that the communities are in order by size.
lay2$community <- the_communities$comm
mcg_mlg_graph <- ggraph(lay2) +
  geom_node_circle(aes(r = size, lty = type, fill = community)) +
  geom_edge_link(aes(start_cap = circle(node1.size, unit = "native"), 
                     end_cap = circle(node2.size, unit = "native"), 
                     width = weight),
                 arrow = arrow(length = unit(0.01, "native"))) +
  geom_node_text(aes(label = gsub("MLG.", "", name), 
                     color = community, 
                     size = size/10), 
                 show.legend = FALSE) +
  coord_fixed() +
  scale_linetype_manual(values = c(1, 3)) +
  scale_edge_width(range = c(0.25, 1.5)) +
  scale_color_manual(values = c("black", "black", "white", "white")) +
  scale_fill_viridis(direction = -1, discrete = TRUE) +
  theme_graph(base_size = 16, base_family = "Helvetica") +
  theme(legend.position = "bottom", legend.direction = "vertical") +
  ggtitle("Relation of Multilocus Genotypes and MCGs")

@@

>>>>

the_communities <- data_frame(vertex = lay2$name, community = lay2$community) %>%
  mutate(comm = case_when(
    .$community == 7 ~ "A",
    .$community == 5 ~ "B",
    .$community == 1 ~ "C",
    TRUE ~ as.character(.$community)
  )) %>%
  group_by(community) %>%
  mutate(comm = ifelse(n() > 10, paste("Community", comm), "Other Communities (n < 10)")) %>%
  mutate(comm = comm) # It just so happens that the communities are in order by size.
lay2$community <- the_communities$comm
mcg_mlg_graph <- ggraph(lay2) +
  geom_node_circle(aes(r = size, lty = type, fill = community)) +
  geom_edge_link(aes(start_cap = circle(node1.size, unit = "native"), 
                     end_cap = circle(node2.size, unit = "native"), 
                     width = weight),
                 arrow = arrow(length = unit(0.01, "native"))) +
  geom_node_text(aes(label = gsub("MLG.", "", name), 
                     color = community, 
                     size = size/10), 
                 show.legend = FALSE) +
  coord_fixed() +
  scale_linetype_manual(values = c(1, 3)) +
  scale_edge_width(range = c(0.25, 1.5)) +
  scale_color_manual(values = c("black", "black", "white", "white")) +
  scale_fill_viridis(direction = -1, discrete = TRUE) +
  theme_graph(base_size = 16, base_family = "Helvetica") +
  theme(legend.position = "bottom", legend.direction = "vertical") +
  ggtitle("Relation of Multilocus Genotypes and MCGs")

@@

## Assessing MCG/MLG community structure/cohesiveness

The communities defined above are defined by the connections between different
MCGs. There is still the question of how cohesive these communities are. I 
believe one way to investigate this is to add the communities assignments back
to the data and assess how well they split up on a dendrogram created with 
genetic distance.

Note in this case we are assigning MLGs to the communities indicated. This means
that even though they share a connection with an MCG in a different community,
they belong to the MCG community they were initially assigned to. We will also
assign communities based on what MCG they belong to.

```{r treebuilding, cache = TRUE}
the_communities
strat_mlg <- strat %>% 
  mutate(MLG = paste0("MLG.", MLG)) %>%
  left_join(the_communities, by = c("MLG" = "vertex"))
strat_mcg <- strat %>% 
  left_join(the_communities, by = c("MCG" = "vertex"))

datdist <- bruvo.dist(dat11, replen = other(dat)$REPLEN)
datree <- phangorn::upgma(datdist)
```

Now we can visualize that tree.

```{r, fig.width = 7, fig.height = 7}
ape::plot.phylo(ape::ladderize(datree), type = "fan", no.margin = FALSE, tip.col = viridis(4, direction = -1)[as.integer(factor(strat_mlg$comm))])
legend(-0.015, 0.11, legend = levels(factor(strat_mlg$comm)), fill = viridis(4, direction = -1), bty = "n")
title("MLG-based assignment")
ape::plot.phylo(ape::ladderize(datree), type = "fan", no.margin = FALSE, tip.col = viridis(4, direction = -1)[as.integer(factor(strat_mcg$comm))])
legend(-0.015, 0.11, legend = levels(factor(strat_mcg$comm)), fill = viridis(4, direction = -1), bty = "n")
title("MCG-based assignment")
```

From what I can see, it appears that there's not much of a tight correlation of 
genetic structure and MCG.

<details>
<summary>Session Information</summary>

```{r, echo = FALSE}
options(width = 100)
devtools::session_info()
```

</details>
